---
title: "R Code - Basic forecasting of average daily temperature in Manaus city, Brazil"
subtitle: "Applied Forecasting course, VNUK Institute for Research and Executive Education"
author: "Hieu Ng. Minh, Long Ng. Hoang"
date: "February 2023"
output: html_notebook
---

Install necessary packages:
```{r include=FALSE}
#install.packages('fpp2')
#install.packages('tidyverse')
#install.packages('GGally')
```
Load packages:
```{r include=FALSE}
#library(fpp2)
#library(tidyverse)
#library(GGally)
```

Read (input) data for training:
```{r}
data_to2018 = read.csv(file = 'data_pre-processed/data_daily_avg_from2000_365_withdate_2000_to_2018.csv')
rownames(data_to2018) = data_to2018$Date
data_to2018$Date = NULL
str(data_to2018)
summary(data_to2018)
```
Convert train data columns to time-series objects:
```{r}
data_to2018_ts = ts(data_to2018, start=c(2000, 1), frequency = 365)
```

Read data for testing:
```{r}
data_2019 = read.csv(file = 'data_pre-processed/data_daily_avg_from2000_365_withdate_2019.csv')
rownames(data_2019) = data_2019$Date
data_2019$Date = NULL
str(data_2019)
summary(data_2019)
```
Convert test data columns to time-series objects:
```{r}
data_2019_ts = ts(data_2019, start=c(2019, 1), frequency = 365)
```


Perform Exploratory Data Analysis (EDA) with the train dataset:

Plot the data:
```{r}
autoplot(data_to2018_ts, facets = T) +
  xlab('Time (from 2000 to 2018)') + ylab('') +
  ggtitle('Daily average weather components from 2000 to 2018 in Manaus city, Brazil')
```

Seasonal plot of temperature only:
```{r}
ggseasonplot(data_to2018_ts[,'Avg_temp'], year.labels=TRUE, year.labels.left=TRUE) +
  ylab("Celsius degree") +
  ggtitle("Seasonal plot: average daily temperature in Manaus, Brazil from 2000 to 2018")
```

Scatterplot matrices to explore the correlation (relationship) between weather components:
```{r}
GGally::ggpairs(data_to2018)
```

Lag plots
```{r}
gglagplot(data_to2018_ts[,'Avg_temp'])
```

Autocorrelation:
```{r}
ggAcf(data_to2018_ts[,'Avg_temp']) +
  ggtitle('ACF graph of average daily temperature in Manaus, Brazil from 2000 to 2018')
```

Data decomposition:

Classical additive decomposition:
```{r}
data_to2018_ts[,'Avg_temp'] %>% decompose(type="additive") %>%
  autoplot() + xlab("Year") +
  ggtitle("Classical additive decomposition of average daily temperature in Manaus")
```

STL decomposition:
```{r}
temp_to2018_tsstl = mstl(data_to2018_ts[,'Avg_temp'])
temp_to2018_tsstl %>% autoplot() +
  ggtitle("STL decomposition of average daily temperature in Manaus")
```
FORECASTING METHODS/MODELS AND RESULTS
```{r}
temp_to2018_ts = data_to2018_ts[,'Avg_temp']
temp_2019_ts = data_2019_ts[,'Avg_temp']
```


Bench marking methods

Average

```{r}
autoplot(temp_to2018_ts) +
  autolayer(temp_2019_ts, series = 'Real data 2019') +
  autolayer(meanf(temp_to2018_ts, h=365), series = 'Mean', PI=F) +
  ggtitle("Forecasting with Average method for 2019 daily average temperature") +
  xlab("Time") + ylab("Celsius degree") +
  guides(colour=guide_legend(title = "Method"))
```
```{r}
checkresiduals(meanf(temp_to2018_ts, h=365))
accuracy(meanf(temp_to2018_ts, h=365))
```


Naive
```{r}
autoplot(temp_to2018_ts) +
  autolayer(temp_2019_ts, series = 'Real data 2019') +
  autolayer(naive(temp_to2018_ts, h=365), series = 'Naive', PI=F) +
  ggtitle("Forecasting with Naive method for 2019 daily average temperature") +
  xlab("Time") + ylab("Celsius degree") +
  guides(colour=guide_legend(title = "Method"))
```
```{r}
checkresiduals(naive(temp_to2018_ts, h=365))
accuracy(naive(temp_to2018_ts, h=365))
```

Seasonal naive
```{r}
autoplot(window(temp_to2018_ts, start=c(2017,1))) +
  autolayer(temp_2019_ts, series = 'Real data 2019') +
  autolayer(snaive(temp_to2018_ts, h=365), series = 'Seasonal naive', PI=F) +
  ggtitle("Forecasting with Seasonal naive method for 2019 daily average temperature") +
  xlab("Time") + ylab("Celsius degree") +
  guides(colour=guide_legend(title = "Method"))
```
```{r}
checkresiduals(snaive(temp_to2018_ts, h=365))
accuracy(snaive(temp_to2018_ts, h=365))
```

Drift
```{r}
autoplot(temp_to2018_ts) +
  autolayer(temp_2019_ts, series = 'Real data 2019') +
  autolayer(rwf(temp_to2018_ts, drift=T, h=365), series = 'Drift', PI=F) +
  ggtitle("Forecasting with Drift method for 2019 daily average temperature") +
  xlab("Time") + ylab("Celsius degree") +
  guides(colour=guide_legend(title = "Method"))
```
```{r}
checkresiduals(rwf(temp_to2018_ts, drift=T, h=365))
accuracy(rwf(temp_to2018_ts, drift=T, h=365))
```

Regression method
```{r}
fit.reg = tslm(Avg_temp ~ Humidity + Atmospheric_pressure + Wind_speed + Cloudiness, data = data_to2018_ts)
fit.reg
```
```{r}
fcast.reg = forecast(fit.reg, newdata = data_2019[,c("Humidity", "Atmospheric_pressure", "Wind_speed", "Cloudiness")])
autoplot(window(temp_to2018_ts, start=c(2017,1))) +
  autolayer(fcast.reg, PI=T, series="Regression model forecast") +
  autolayer(temp_2019_ts, series="Real 2019 data") +
  guides(colour=guide_legend(title="")) +
  ggtitle("Forecasting with Regression model for 2019 daily average temperature") +
  ylab("Celsius degree")
```
```{r}
checkresiduals(fcast.reg)
accuracy(fcast.reg)
```
```{r}
fcast.reg = forecast(fit.reg, newdata = data_2019[,c("Humidity", "Atmospheric_pressure", "Wind_speed", "Cloudiness")])
autoplot(temp_to2018_ts) +
  autolayer(fcast.reg, PI=T, series="Regression model forecast") +
  autolayer(temp_2019_ts, series="Real 2019 data") +
  guides(colour=guide_legend(title="")) +
  ggtitle("Forecasting with Regression model for 2019 daily average temperature (full)") +
  ylab("Celsius degree")
```



Decomposition with STL
```{r}
fcast.stl_ets = stlf(temp_to2018_ts, h=365)
autoplot(window(temp_to2018_ts, start=c(2017,1))) +
  autolayer(fcast.stl, PI=T, series="STL decomposition forecast") +
  autolayer(temp_2019_ts, series="Real 2019 data") +
  guides(colour=guide_legend(title="")) +
  ggtitle("Forecasting with STL + ETS approach for 2019 daily average temperature") +
  ylab("Celsius degree")
```

```{r}
checkresiduals(fcast.stl_ets)
accuracy(fcast.stl_ets)
```

```{r}
fcast.stl = stlf(temp_to2018_ts, method='naive', h=365)
autoplot(window(temp_to2018_ts, start=c(2017,1))) +
  autolayer(fcast.stl, PI=T, series="STL decomposition forecast") +
  autolayer(temp_2019_ts, series="Real 2019 data") +
  guides(colour=guide_legend(title="")) +
  ggtitle("Forecasting with STL + Random walk for 2019 daily average temperature") +
  ylab("Celsius degree")
```

```{r}
checkresiduals(fcast.stl)
accuracy(fcast.stl)
```

ETS
```{r}
fit.ets = ets(temp_to2018_ts)
summary(fit.ets)
```

```{r}
autoplot(fit.ets)
```

```{r}
autoplot(window(temp_to2018_ts, start=c(2017,1))) +
  autolayer(forecast(fit.ets, h=365), PI=T, series="ETS(M,N,N) forecast") +
  autolayer(temp_2019_ts, series="Real 2019 data") +
  guides(colour=guide_legend(title="")) +
  ggtitle("Forecasting with ETS(M,N,N) for 2019 daily average temperature") +
  ylab("Celsius degree")
```

```{r}
checkresiduals(forecast(fit.ets, h=365))
accuracy(forecast(fit.ets, h=365))
```
Holt-Winters' seasonal method
```{r}
fit.hw = hw(temp_to2018_ts, seasonal="additive")
fit.hw
```
Holt-Winters' damped method
```{r}
fit.hw = hw(temp_to2018_ts, damped=T, seasonal="additive")
fit.hw
```


ARIMA

```{r}
fit.arima = auto.arima(temp_to2018_ts)
fit.arima
```

```{r}
autoplot(window(temp_to2018_ts, start=c(2017,1))) +
  autolayer(forecast(fit.arima, h=365), PI=T, series="ARIMA(3,1,3) forecast") +
  autolayer(temp_2019_ts, series="Real 2019 data") +
  guides(colour=guide_legend(title="")) +
  ggtitle("Forecasting with ARIMA(3,1,3) for 2019 daily average temperature") +
  ylab("Celsius degree")
```

```{r}
checkresiduals(forecast(fit.arima, h=365))
accuracy(forecast(fit.arima, h=365))
```


Dynamic regression model with ARIMA errors
```{r}
fit.dreg = auto.arima(temp_to2018_ts, xreg=data_to2018_ts[,c("Humidity", "Atmospheric_pressure", "Wind_speed", "Cloudiness")])
fit.dreg
```

```{r}
fcast.dreg = forecast(fit.dreg, xreg=data_2019_ts[,c("Humidity", "Atmospheric_pressure", "Wind_speed", "Cloudiness")])
autoplot(window(temp_to2018_ts, start=c(2017,1))) +
  autolayer(fcast.dreg, PI=T, series="ARIMA(5,1,1) errors") +
  autolayer(temp_2019_ts, series="Real 2019 data") +
  guides(colour=guide_legend(title="")) +
  ggtitle("Forecasting with Dynamic Regression model with ARIMA errors for 2019") +
  ylab("Celsius degree")
```

```{r}
checkresiduals(fcast.dreg)
accuracy(fcast.dreg)
```

Dynamic harmonic regression
```{r}
#install.packages("gridExtra")
#library(gridExtra)
```


```{r}
plots <- list()

for (i in seq(6)) {
  fit <- auto.arima(temp_to2018_ts, xreg = fourier(temp_to2018_ts, K = i),
    seasonal = FALSE, lambda = 0)
  
  plots[[i]] <- autoplot(forecast(fit,
      xreg=fourier(temp_to2018_ts, K=i, h=365))) +
    xlab(paste("K=",i,"   AICC=",round(fit[["aicc"]],2))) +
    ylab("") +
    ggtitle(as.character(fit))
}
gridExtra::grid.arrange(
  plots[[1]],plots[[2]],plots[[3]],
  plots[[4]],plots[[5]],plots[[6]], nrow=3)
```
